---
title: "Explore COVID-19 pandemic with R"
description: |
  A sneak peak into the COVID-19 data from Africa.
author:
  - name: Stephen Balogun
    url: https://taiye.netlify.app
date: 2022-01-11
output:
  distill::distill_article:
    self_contained: false
categories:
  - health
---


```{r include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = NA, tidy = FALSE, echo = FALSE)
source("./covid19.R")
```

## Introduction 

It is no longer news that COVID-19 has witnessed exponential growth since the beginning of March 2020, spreading from **`r nrow(n_count)`** countries at the end of February 2020 to **`r nrow(n_count2)`** as at the end of 2021. The number of of confirmed cases have also grown astronomically from *86, 000* to hundreds of millions during the same period of time. We briefly look at this global trend and compare this with the current trend seen in Africa. The trend in Africa is much similar to the global trend, except for some time lag in the onset of rise with less than 5 confirmed cases at the beginning March 2020 to over one million cases confirmed cases within a period of 6 months.   


## Trend  {.tabset .tabset-fade .tabset-pills}


<!-- ### Globe -->

```{r}
ggplotly(global_trend)
```


<!-- ### Africa -->


```{r}
ggplotly(africa_trend)
```

<!-- ## Spread across countries -->

```{r}
ggplotly(n_countries)
```



## Country-specific Trends

It is no longer news that China, being the country where the first case was discovered experienced the pandemic much early and appear to have been able to contain the disease. The other country that has done remarkably well in containment appears to be South Korea. The country was able to flattening the curve early, something most other countries experiencing exponential growth have not been able to do.


<!-- ### Globe -->

```{r}
ggplotly(world)
```

### Top 10 countries



```{r}

ggplotly(world2_plot)

```



### Africa


```{r}

ggplotly(africa2_plot)

```


```{r}

ggplotly(world2_plot)

```



## Nigeria

Nigeria recorded its first case of covid-19 on the 28th of February 2020 and did not have a second case till a week after (on the 9th of March 2020). However, within the space of another 8 days, the number of cases jumped from **2** (on the 16th of March 2020 to **30** cases by 22nd March 2020). Since then, there has been 3 demonstrable waves of rapid increase infection rates. There appears to be a third wave, starting in December 2021, coinciding with the yuletide and the rapid spread of the Omicron variant. 

In another discussion, we will examine how the vaccination rates have affected incidence and mortality from COVID-19. Meanwhile, we all owe it to one another to continue to abide by recommended safety guidelines provided by [**NCDC**](http://covid19.ncdc.gov.ng/faq.html#about), [**WHO**](https://www.who.int/emergencies/diseases/novel-coronavirus-2019/advice-for-public), [**CDC**](https://www.cdc.gov/coronavirus/2019-ncov/index.html), and other reputable sites for more information.  


```{r}
ggplotly(ngr)

```


## Conclusion

The world has demonstrated that panic and fear only fuels the pandemic, there is a need for us to continue to be rational and observe the safety measures advised. In so doing, we will not only be protecting ourselves, but also our families and loved ones, and the general community.


```{r}
ggplotly(animate)

```

